---
title: "Variables Codebook"
subtitle: "Data cleaning project JHU"

author: "Federica Gazzelloni"
date: "23/12/2020"

output:
  html_document:
    toc: TRUE
    number_sections: FALSE
    toc_float:
      collapsed: TRUE
      smooth_scroll: FALSE
      
    keep_md: yes
---


```{r}
library(rmarkdown)
library(knitr)
```


# Project Description
It is a randomized sample training and test analysis of a group of 30 volunteers within an age bracket of 19-48 years.

Individuals performed six activities:

- WALKING
- WALKING_UPSTAIRS
- WALKING_DOWNSTAIRS
- SITTING
- STANDING
- LAYING

wearing a smartphone (Samsung Galaxy S II) on the waist to capture acceleration and velocity.



# Study design and data processing 

### Randomize experiment divided the group into:{.tabset}

- training 70%
- test 30%

### Variation
The variation measured through sensor signals captured by "t" time-domain and "-XYZ" 3-axial signals in the X, Y and Z directions, indicated as:

### 
* "tAcc-XYZ"  accelerometer subgroupped into:
                               
                               >- "tBodyAcc-XYZ"      body acceleration 
                               >- "tGravityAcc-XYZ"   gravity
* "tGyro-XYZ" gyroscope

In particular, signals were captured at a constant of 50Hz and then filtered by:
- median
- third order removal noise corner frequency of 20 Hz
- subgroupped corner frequency of 0.3 Hz


A list of variables taken into consideration for the measurement of the data processing are subdivided by the nature, magnitude and frequency, as described below.


### Jerk signals
Jerk signals obtained deriving the body linear acceleration and angular velocity:
- tBodyAccJerk-XYZ
- tBodyGyroJerk-XYZ

### Magnitude
The magnitude of these three-dimensional signals were calculated using the Euclidean norm:
- tBodyAccMag
- tGravityAccMag

- tBodyGyroMag

- tBodyAccJerkMag
- tBodyGyroJerkMag

### Frequency
The Fast Fourier Transform (FFT) was applied to some of these signals producing,"f" frequency domain signals:
- fBodyAcc-XYZ 
- fBodyGyro-XYZ

- fBodyAccJerk-XYZ
- fBodyAccJerkMag

- fBodyGyroMag
- fBodyGyroJerkMag

### Descriptive statistics
Finally,descriptive statistics were calculated to analyze results,in particular the following are the two measures of interest in this project:  
- mean(): Mean value
- std(): Standard deviation


### Collection of the raw data
The experiments have been video-recorded to label the data manually, and randomly partitioned into two sets.



# Creating the tidy datafile

### Guide to create the tidy data file
1. download the data and unzip it in the project folder
2. check of the original "Readme.txt" and "features_info.txt" files for information about how variables are listed and what is their meaning
3. identify the location of the variables of interest for the project
4. check the structure of the data sets for merging column names with content
5. extracting needed information by the selection of the columns containing the descriptive statistics as requested.
6. split the columns to selected information to obtain a tidy data set


### Cleaning of the data
The cleaning script download and unzip the data sets, to check the requested information and the structure of the sets for merging them in a unique dataframe.

```{r}
#source("README.md")
```



# Description of the variables
General description of the file including:
 - Dimensions of the dataset
 - Summary of the data
 - Variables present in the dataset

(you can easily use Rcode for this, just load the dataset and provide the information directly form the tidy data file)

### Variable 1 (repeat this section for all variables in the dataset)
Short description of what the variable describes.

Some information on the variable including:
 - Class of the variable
 - Unique values/levels of the variable
 - Unit of measurement (if no unit of measurement list this as well)
 - In case names follow some schema, describe how entries were constructed (for example time-body-gyroscope-z has 4 levels of descriptors. Describe these 4 levels). 

(you can easily use Rcode for this, just load the dataset and provide the information directly form the tidy data file)

#### Notes on variable 1:
If available, some additional notes on the variable not covered elsewehere. If no notes are present leave this section out.

# Sources
Sources you used if any, otherise leave out.

## Annex
If you used any code in the codebook that had the echo=FALSE attribute post this here (make sure you set the results parameter to 'hide' as you do not want the results to show again)

```{r}
par(mar = c(4, 4, 0.1, 0.1))
plot(pressure, type = "b", pch = 19)
```

```{r}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```



















